---
title: "assignment"
author: "Kiran Kumar Patro"
date: "July 23, 2017"
output: html_document
---


#1. Histograms for "Age", "CCAvg", "Experience".
```{r}
library(ggplot2)

Universal_Bank <- read.csv("UniversalBank.csv")
colnames(Universal_Bank)
str(Universal_Bank)
sum(is.na(Universal_Bank))
summary(Universal_Bank)


```


```{r}
attach(Universal_Bank)

default_par <- par()
par(mfrow=c(1,3))

par(bg="grey")
hist(Age, col="brown", xlab="Average",
     main="Frequency Histogram:Average",xlim = c(20,70), ylim = c(0,700))
hist(CCAvg, col="red", xlab="CCAvg",
     main="Frequency Histogram:CCAvg", xlim = c(0,10), ylim = c(0,700)) 
hist(Experience, col="red", xlab="Experience",
     main="Frequency Histogram:Experience", xlim = c(-5,50), ylim = c(0,700))

```

#2. Boxplots for "Income". Add the mean line to the plots.

```{r}
par(mfrow=c(1,1))
boxplot(Income,main = "Box Plot: Income", ylab = "Income")
```


#3. Boxplots for "Income" w.r.t. "Education"

```{r}
par(mfrow=c(1,1))
boxplot(Income ~ Education, 
        main = "Income Vs. Education", xlab="Education", ylab="Income")
```

#4. Bar-plots for "Personal Loan". Add color, titles, labels etc.

```{r}
ggplot(universalbank_data,aes(x=Personal.Loan)) + geom_bar() + 
  xlab("Personal loan") + ylab("ID")
ggplot(universalbank_data,aes(x=Personal.Loan)) + geom_bar() +
  xlab("Personal Loan") + ylab("# ID")  +
  ggtitle("#Personal loan")+
  geom_text(stat='count',aes(label=..count..),vjust=0)
```

#5. Bar-plots for "Education"; play with bar width, background, text, legend etc.

```{r}
ggplot(universalbank_data,aes(x=Education, fill=Education)) + geom_bar() +
  xlab("Education") + ylab("# ID")  +
  ggtitle("#Education")
```


#6. Scatter plots for Age vs. Income, Experience vs. Income

```{r}
ggplot(data=universalbank_data, aes(x=Age, y=Income))+ geom_point() + 
  xlab("Age")+ylab("Income") + 
  ggtitle("Age vs. Income")
```

```{r}
ggplot(data=universalbank_data, aes(x=Experience, y=Income))+ geom_point() + 
  xlab("Experience")+ylab("Income") + 
  ggtitle("Experience vs. Income")
```


#7. Scatter plots for Age vs. Income. Also add whether or not the customer took the personal loan (hint: can use color, shape etc. as the 3rd dimension)

```{r}
ggplot(data=universalbank_data, aes(x=Age, y=Income, colour = grey.colors(1,start=0.3,end = 0.9, gamma = 2.2,alpha = NULL) + 
  geom_point() + 
  xlab("Age")+ylab("Income") + 
  ggtitle("Age vs. Income")
```

#Plot Education Vs. Income. Add the following to the plot: whether or not the customer accepted the personal loan, CCAvg.

```{r}
 ggplot(data=universalbank_data, aes(x=Education, y=Income))+ geom_point() + 
    xlab("Education")+ylab("Income") + 
    ggtitle("Education vs. Income")
```





